The title

### 2021-08-25 manuscript/manuscript.Rmd
Add TFR analysis and plot

  The authors declare no competing financial interests.
  The preprocessed data and analaysis code for this study are openly available at https://osf.io/uksbc/.
abstract: |
  Does our perception of an object change as soon as we discover what function it serves?
  This question matters not only when we encounter novel tools or gadgets in our everyday lives, but also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities.
  Here, we showed human participants (*n* = 48) pictures of unfamiliar objects either together with words matching their function, leading to semantically informed perception, or together with non-matching words, resulting in naive perception.
  We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
  We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component (150--200 ms) and reduced amplitudes in the N400 component (400--700 ms).
  When the same objects were presented once more without any information, the N400 effect persisted and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
  Consistent with previous work, our results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
  We demonstrate for the first time that these effects are instantaneous in the sense that they can be observed within the same trial as well as one trial after the information has been given, providing strong evidence that cognition affects perception.
keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
bibliography      : ["references.bib", "r-references.bib"]
csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man"
fontsize          : "10pt"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
  papaja::apa6_docx: default
editor_options: 
  chunk_output_type: console
header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load packages
library(buildmer)
library(cowplot)
library(emmeans)
library(here)
library(knitr)
library(lme4)
library(lmerTest)
library(magick)
library(MASS)
library(papaja)
library(parallel)
library(reticulate)
library(scales)
library(tidyverse)
library(magrittr)

@@

# Re-run preprocessing and/or model fitting?
run <- list(preprocessing = TRUE, model_fitting = TRUE)

# File names for easier input/output
deps <- c("logRT", "P1", "N170", "N400")  # dependent variables
io <- list(
  montage = "results/montage.csv",
  obj_per_cond = "results/obj-per-cond.csv",
  trials = "results/trials.csv",
  ave_grand = "results/ave_grand.csv",
  ave_tfr_flat = "results/ave-tfr_flat.csv",
  ave_tfr_diff = "results/ave-tfr_diff.csv",
  models_lme = "results/models_lme.RDS",
  models_brm = str_c("results/model_brm_", deps, ".RDS") %>% set_names(deps),
  model_brm_prior = "results/model_brm_prior.RDS",
  models_control = "results/models_control.RDS"
)
```

@@

```{r, fig3, eval=FALSE, include=TRUE}
# Read grand averages of evoked power
ave_tfr_flat <- read_csv(io$ave_tfr_flat, col_types = cols(), progress = FALSE)
ave_tfr_diff <- read_csv(io$ave_tfr_diff, col_types = cols(), progress = FALSE)

# Scaling of power in all plots
limits_power <- c(0, 0.5)

# Time-frequency of interest based on visual inspect
tfoi <- list(tmin = 150, tmax = 350, fmin = 14, fmax = 24) # Beta or
tfoi <- list(tmin = 200, tmax = 400, fmin = 34, fmax = 48) # Gamma

# Standard text format
standard_text <- element_text(size = 10, color = "black", family = "Helvetica")

plot_tfr_raster <- function(ave_tfr, limits, tmin, tmax, fmin, fmax, label) {
  # Average across electrodes
  ave_tfr %>%
    pivot_longer(-c(part, time, freq), names_to = "electrode", values_to = "power") %>%
    group_by(part, time, freq) %>%
    summarise(power = mean(power), .groups = "drop") %>%
    # Create one raster plot per part
    group_by(part) %>%
    group_map(function(data_part, part) {
      ggplot(data_part, aes(x = time, y = freq, fill = power)) +
        geom_raster() +
        scale_fill_viridis_c(
          limits = limits,
          labels = function(x) x * 100,
          guide = guide_colorbar(
            title = "Power (% change\nfrom baseline)",
            title.position = "left",
            title.theme = element_text(size = 10, color = "black", angle = 90, lineheight = 1),
            title.hjust = 0.5,
            label.theme = element_text(size = 10, color = "black"),
            barheight = unit(4, "cm"),
            ticks = FALSE,
            direction = "vertical"
          ),
          oob = squish
        ) +
        scale_x_continuous(breaks = seq(-100, 700, 200)) +
        coord_cartesian(expand = FALSE) +
        annotate("rect", xmin = tmin, xmax = tmax, ymin = fmin + 1, ymax = fmax + 1, color = "white", fill = NA, linetype = "dotted") +
        annotate("segment", x = 0, xend = 0, y = -Inf, yend = Inf, color = "white", linetype = "dotted") +
        labs(x = "Time (ms)", y = "") +
        theme_minimal() +
        theme(
          legend.position = "none",
          axis.title = standard_text,
          axis.text = standard_text
        )
    }) -> plotlist
  # Add label and axis title to the first plot
  plotlist[[1]] <- plotlist[[1]] + annotate(
    "text",
    x = -100,
    y = 27,
    label = label,
    size = 10 / .pt,
    angle = 90,
    family = "Helvetica",
    color = "white",
    lineheight = 0.9
  ) +
    ylab("Frequency (ms)")
  # Extract legend for another subplot
  current_legend <<- get_legend(plotlist[[1]] + theme(legend.position = "right"))
  plot_grid(plotlist = plotlist, nrow = 1)
}

plot_tfr_topo <- function(ave_tfr, limits, tmin, tmax, fmin, fmax, legend) {
  # Average across times and frequencies of interest
  ave_tfr_diff %>%
    # Select time window and frequencies of interest
    filter(time >= tmin & time <= tmax & freq >= fmin & freq <= fmax) %>%
    # Convert to long format
    pivot_longer(-c(part, time, freq), names_to = "electrode", values_to = "power") %>%
    # Create one topoplot per part
    group_by(part) %>%
    group_map(function(data_part, part) {
      data_part %>%
        group_by(electrode) %>%
        summarise(power = mean(power), .groups = "drop") %>%
        inner_join(montage, by = "electrode") %>%
        eegUtils::topoplot(
          limits = limits,
          palette = "viridis",
          contour = FALSE,
          quantity = "power",
          scaling = 0.5
        ) +
        theme(legend.position = "none", )
    }) -> plotlist
  # plotlist <- c(plotlist, list(legend))
  plot_grid(
    NULL,
    plotlist[[1]],
    NULL,
    plotlist[[2]],
    NULL,
    plotlist[[3]],
    nrow = 1,
    rel_widths = c(2, 7, 0, 7, 1, 7)
  )
}

# Bring it all together
plot_grid(
  plot_headings(spacing = c(1, 7, 1, 7, 1, 7)),
  do.call(
    plot_tfr_raster, c(tfoi, list(
      ave_tfr = ave_tfr_flat,
      limits = limits_power,
      label = "All channels,\nall conditions"
    ))
  ),
  do.call(
    plot_tfr_raster, c(tfoi, list(
      ave_tfr = ave_tfr_diff,
      limits = limits_power,
      label = "All channels,\n\u0394 (informed - naive)"
    ))
  ),
  do.call(
    plot_tfr_topo, c(tfoi, list(
      ave_tfr = ave_tfr_diff,
      limits = limits_power,
      legend = current_legend
    ))
  ),
  nrow = 4,
  rel_heights = c(1, 5, 5, 5),
  labels = c("", "A", "B", "C"),
  label_y = c(1, 1.16, 1.1, 1),
  label_fontfamily = "Helvetica"
) +
  draw_plot(current_legend, x = -0.42, y = -0.34) +
  annotate_line(0.344, 0.344, -Inf, Inf) +
  annotate_line(0.676, 0.676, -Inf, Inf) +
  ggsave("data/processed/tfr.pdf", width = unit(10, "cm"), height = unit(7, "cm"), device = cairo_pdf)
```

@@

>>>>

  The authors declare no competing financial interests.
  The preprocessed data and analaysis code for this study are openly available at https://osf.io/uksbc/.
abstract: |
  Does our perception of an object change as soon as we discover what function it serves?
  This question matters not only when we encounter novel tools or gadgets in our everyday lives, but also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities.
  Here, we showed human participants (*n* = 48) pictures of unfamiliar objects either together with words matching their function, leading to semantically informed perception, or together with non-matching words, resulting in naive perception.
  We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
  We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component (150--200 ms) and reduced amplitudes in the N400 component (400--700 ms).
  When the same objects were presented once more without any information, the N400 effect persisted and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
  Consistent with previous work, our results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
  We demonstrate for the first time that these effects are instantaneous in the sense that they can be observed within the same trial as well as one trial after the information has been given, providing strong evidence that cognition affects perception.
keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
bibliography      : ["references.bib", "r-references.bib"]
csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
  papaja::apa6_docx: default
editor_options: 
  chunk_output_type: console
header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load packages
library(buildmer)
library(cowplot)
library(emmeans)
library(here)
library(jsonlite)
library(knitr)
library(lme4)
library(lmerTest)
library(magick)
library(MASS)
library(papaja)
library(parallel)
library(reticulate)
library(scales)
library(tidyverse)
library(magrittr)

@@

# Re-run preprocessing and/or model fitting?
run <- list(
  preprocessing = FALSE,
  model_fitting = FALSE,
  time_frequency = FALSE
)

# File names for easier input/output
deps <- c("logRT", "P1", "N170", "N400") # dependent variables
io <- list(
  montage = "results/montage.csv",
  obj_per_cond = "results/obj-per-cond.csv",
  trials = "results/trials.csv",
  ave_grand = "results/ave_grand.csv",
  ave_tfr_flat = "results/ave-tfr_flat.csv",
  ave_tfr_diff = "results/ave-tfr_diff.csv",
  models_lme = "results/models_lme.RDS",
  models_brm = str_c("results/model_brm_", deps, ".RDS") %>% set_names(deps),
  model_brm_prior = "results/model_brm_prior.RDS",
  models_control = "results/models_control.RDS",
  tfr_stats = "results/tfr_stats.json",
  tfr_grand_ave = "results/tfr_grand_ave.csv"
)
```

@@

@@

## Time-frequency analysis

```{python, tfr, eval=run$time_frequency}
import numpy as np
import pandas as pd
from mne import read_epochs
from mne.channels import find_ch_adjacency
from mne.stats import combine_adjacency, permutation_cluster_1samp_test
from scipy import sparse

# Read evoked power for all subjects
fnames = [f"data/preprocessed/{sid}_tfr-ave.csv" for sid in r.subject_id]
tfrs = [pd.read_csv(fname) for fname in fnames]
tfrs = pd.concat(tfrs)

# Get electrode names
electrodes = r.montage["electrode"].tolist()

# Define non-channel columns to sort by
idx_cols = ["condition", "subject_id", "time", "freq"]

# Prepare empty lists to store outputs of permutation tests
t_obs = []
clusters = []
p_values = []
h0_max = []

# Loop over parts
parts = ["I", "II", "III"]
for part in parts:

    # Get data for the current part
    df = tfrs[tfrs["part"] == part]
    df = df.sort_values(idx_cols)

    # Extract TFR data per condition
    dat_informed = df.loc[df["condition"] == "Informed", electrodes].values
    dat_naive = df.loc[df["condition"] == "Naive", electrodes].values
    dat_diff = dat_informed - dat_naive  # Shape (subs x times x freqs) x elecs

    # Reshape to X matrix for permutation test
    n_subjects = len(df["subject_id"].unique())
    n_times = len(df["time"].unique())
    n_freqs = len(df["freq"].unique())
    X = dat_diff.reshape(n_subjects, n_times, n_freqs, -1)

    # # Plot the data matrix, averaged across participants and channels
    # import matplotlib.pyplot as plt
    # plt.imshow(X.mean(axis=(0, 3)).T, aspect="auto", vmin=-0.5, vmax=0.5, origin='lower'); plt.show()

    # Define adjacency matrix for frequencies
    adjacency_freq = sparse.csr_matrix(np.ones((n_freqs, n_freqs)))

    # Define adjacency matrix for channels based on channel positions
    example_epochs = read_epochs("data/preprocessed/sub-01_epo.fif", preload=False)
    adjacency_chan, ch_names = find_ch_adjacency(example_epochs.info, ch_type="eeg")

    # Combine adjacency matrices for frequencies and channels
    adjacency = combine_adjacency(adjacency_freq, adjacency_chan)

    # # Plot the combined adjacency matrix
    # plt.imshow(adjacency.toarray(), origin='lower', interpolation='nearest'); plt.show()

    # Run permutation test and save results
    cluster_stats = permutation_cluster_1samp_test(
        X, adjacency=adjacency, n_permutations=1000
    )

    # Store outputs
    t_obs.append(cluster_stats[0])
    clusters.append(cluster_stats[1])
    p_values.append(cluster_stats[2])
    h0_max.append(cluster_stats[3])

# Gather outputs in DataFrame and save
cluster_stats_df = pd.DataFrame(
    {
        "part": parts,
        "t_obs": t_obs,
        "clusters": clusters,
        "p_values": p_values,
        "h0_max": h0_max,
    }
)
cluster_stats_df.to_json(r.io["tfr_stats"])

# Export averaged data for plotting
ave_cols = ["part", "condition", "time", "freq"]
tfrs_plotting = tfrs.groupby(ave_cols, as_index=False).mean()  # Across subjects
tfrs_plotting["power"] = tfrs_plotting[electrodes].mean(axis=1)  # Across channels
tfrs_plotting = tfrs_plotting.drop(electrodes, axis=1)
tfrs_plotting.to_csv(r.io["tfr_grand_ave"], index=False)
```

```{r, fig2, include=TRUE, fig.height=4, fig.cap="(ref:figure-2-caption)"}
# Load permutation test results
jsonlite::read_json(io$tfr_stats) %>%
  as_tibble() -> tfr_stats

# Get minimum cluster p value per part
tfr_stats$p_values %>%
  map(as.numeric) %>%
  map(min) %>%
  set_names(c("I", "II", "III")) -> tfr_min_ps

# Plot time-frequency results per part and condition
read_csv(io$tfr_grand_ave, col_types = cols(), progress = FALSE) %>%
  plot_fig2()
```

(ref:figure-2-caption) Looks great!
